---
title: "R codebase for Li, Mac Aogáin & Xu et al. 2022 - Figure 1 (b) and (c)"
authors: "Micheál Mac Aogáin, Hong Sheng Cheng"
date: "4/15/2022"
output: 
  html_document
---

## R Markdown
This R Markdown document details R-generated analysis represented in Figure 1 of "Neissieria species as pathobioints in bronchiectasis" (under review - Cell Host & Microbe).

Requires: DESeq2, GO.db, WGCNA, impute and preprocessCore.
```{r, warning = FALSE, message = FALSE}
if(!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
#install.packages("tidyverse")
#install.packages("vegan")
#install.packages("ggplot2")
#install.packages("Hmisc")
#install.packages("ggpubr")
library(tidyverse)
library(vegan)	
library(ggplot2)	
library(Hmisc)
library(ggpubr)
```

## Load data for Figure 1
```{r}
Master=read.table("../Data/microbiome.csv", header=T,	row.names=1,sep=",")
```

## Plotting Principal component analysis (PCA) human airway microbiome profiles of bronchiectasis patients from SG-KL and DD cohorts
```{r}
Mbiome<-Master %>%
  select(33:166)
  
vegdist(Mbiome, "bray")-> Mbiome_PCoA
as.matrix(Mbiome_PCoA)->Mbiome_PCoA
BrayCurtMbiome=cmdscale(Mbiome_PCoA)

#extract co-ordinates for plotting
BCords<-as.data.frame(scores(BrayCurtMbiome))
Master$BacBC_1<-BCords$Dim1
Master$BacBC_2<-BCords$Dim2

#Plot co-ordinates and ordinate by geographic location
ggplot(data = Master, aes(BacBC_1, BacBC_2, colour = Asian_European, shape = Matching_neisseria))+ 
  geom_point(size =3)+
  scale_colour_manual(values = c("Asian" = 'Red', "European" = 'Blue'), name = "Geographic Origin", labels = c("SG-KL","DD"))+
  scale_shape_manual(values=c(16, 1))+
  theme_set(theme_gray())+
  theme(legend.justification = c(0, 1),legend.position = c(0, 1),
        legend.box.margin=margin(c(4,4,4,4)),
        legend.background = element_blank())+
  labs(
    x = "PC1 (20.3%)", y = "PC2 (15.4%)")+
  labs(size = "% Neisseria spp. \n (relative abundance)")+
  labs(alpha = "% Neisseria spp. \n (relative abundance)")+
  guides(shape = "none")+
  scale_y_reverse()+
  scale_x_reverse()

ggplot(data = Master, aes(BacBC_1, BacBC_2, colour = Neisseria))+ 
  geom_point(size =3)+
  scale_color_gradient(low="grey", high="red", limits = c(0,100))+
  theme_set(theme_gray())+
  theme(legend.justification = c(0, 1),legend.position = c(0, 1),
        legend.box.margin=margin(c(5,5,5,5)),
        legend.background = element_blank())+
  labs(
    x = "PC1 (20.3%)", y = "PC2 (15.4%)")+
  labs(colour = "Neisseria spp. \nrelative abundance (%)")+
  scale_y_reverse()+
  scale_x_reverse()

```

#Boxplot of Neisseria abundance
```{r}
ggplot(data = Master, aes(Asian_European,Neisseria,  
                          group = Asian_European, colour = Asian_European))+ geom_boxplot(outlier.shape = NA, )+
  geom_jitter(width=0.2)+
  scale_colour_manual(values = c("Asian" = 'Red', "European" = 'Blue'), name = "Geographic Origin", labels = c("SG-KL","DD"))+
  stat_boxplot(geom = 'errorbar', width = 0.5)+
  #theme_classic()+
  theme_set(theme_gray())+
  theme(legend.justification = c(0, 1),legend.position = c(0, 1),
        legend.box.margin=margin(c(5,5,5,5)),
        legend.background = element_blank())+
  labs(colour="", 
       x = "", y = "Neisseria spp. relative abundance (%)")+
  scale_x_discrete(
                   labels=c("SG-KL", "DD"))+
  guides(colour = FALSE)

#Test difference in Neisseria abundance  
wilcox.test(Master$Neisseria~Master$Asian_European)

```

rmarkdown::render(rstudioapi::getSourceEditorContext()$path)

