---
title: "R codebase for Li, Mac Aogáin & Xu et al. 2022 - Figure 4 (b) and (c)"
authors: "Micheál Mac Aogáin, Hong Sheng Cheng"
date: "4/18/2022"
output: html_document
---

## R Markdown
This R Markdown document details R-generated analysis represented in Figure 4 of "Neissieria species as pathobioints in bronchiectasis" (under review - Cell Host & Microbe).

## Install and load required R packages
Requires: DESeq2, GO.db, WGCNA, impute and preprocessCore.
```{r, warning = FALSE, message = FALSE}
#if(!require("BiocManager", quietly = TRUE))
#  install.packages("BiocManager")
#BiocManager::install("DESeq2", force = TRUE)
#BiocManager::install("GO.db", force = TRUE)
#BiocManager::install("WGCNA", force = TRUE)
#BiocManager::install("impute", force = TRUE)
#BiocManager::install("preprocessCore", force = TRUE)
library(DESeq2)
library(WGCNA)
library(ggplot2)
library(RColorBrewer)
```

## Load data for Figure 4
```{r}
FC_count<-read.csv("../Data/FC_count.csv")
coldata<-read.csv("../Data/coldata.csv", head=T,row.names=1,stringsAsFactors = FALSE)

#preparing count file from Featurecounts
rownames(FC_count)<-FC_count$gene_id
FC_count$gene_id<-NULL
FC_count<-as.matrix(FC_count)

#preparing coldata
rownames(coldata)<-coldata$ID
coldata$ID<-NULL
coldata$group<-paste0(coldata$Treatment,sep="_D",coldata$Day)
coldata$group<-factor(coldata$group,levels=c("Control_D0","Infected_D2","Infected_D4","Infected_D7","Infected_D14"))
coldata$Treatment<-factor(coldata$Treatment,levels=c("Control","Infected"))
coldata$Day<-factor(coldata$Day,levels=c("0",'2','4','7','14'))

```

## DESeq2 analysis
```{r}
dds<-DESeqDataSetFromMatrix(countData=FC_count,colData = coldata, design=~group)
dds<-DESeq(dds)
vsd<-varianceStabilizingTransformation(dds, blind=TRUE)
```

## Plotting Principal component analysis (PCA) of mouse lung transcriptomes across the 14-day course of infection in Figure 4 (b).
```{r plot panel B}
plotPCA(vsd,intgroup="group")
```

## Data wrangling of DESeq2 outputs: obtain DEGs at D2,D4,D7 & D14
```{r DEGS}
res.D2<-results(dds,contrast=c('group','Infected_D2','Control_D0'),alpha=0.05)
res.D2<-res.D2[order(res.D2$padj),]
sig.res.D2<-res.D2[which((res.D2$padj<0.05) & (abs(res.D2$log2FoldChange)>1)),]

res.D4<-results(dds,contrast=c('group','Infected_D4','Control_D0'),alpha=0.05)
res.D4<-res.D4[order(res.D4$padj),]
sig.res.D4<-res.D4[which((res.D4$padj<0.05) & (abs(res.D4$log2FoldChange)>1)),]

res.D7<-results(dds,contrast=c('group','Infected_D7','Control_D0'),alpha=0.05)
res.D7<-res.D7[order(res.D7$padj),]
sig.res.D7<-res.D7[which((res.D7$padj<0.05) & (abs(res.D7$log2FoldChange)>1)),]

res.D14<-results(dds,contrast=c('group','Infected_D14','Control_D0'),alpha=0.05)
res.D14<-res.D14[order(res.D14$padj),]
sig.res.D14<-res.D14[which((res.D14$padj<0.05) & (abs(res.D14$log2FoldChange)>1)),]

DEGlist<-union(rownames(sig.res.D2),row.names(sig.res.D4))
DEGlist<-union(DEGlist,row.names(sig.res.D7))
DEGlist<-union(DEGlist,row.names(sig.res.D14))

vsdtable<-as.data.frame(assay(vsd))
vsdtable<-vsdtable[,c(1:5,11:24,6:10)]
vsdtable_DEGlist<-subset(vsdtable,rownames(vsdtable) %in% DEGlist)
vsdmat<-as.matrix(vsdtable_DEGlist)
write.csv(vsdmat, "../Data/vsdmat.csv")

coldata_hm<-coldata[c(1:5,11:24,6:10),]
coldata_hm<-data.frame(coldata_hm[,c(3),drop=FALSE])

```

## Plotting heatmap of functional gene clusters from mouse lung transcriptomes in Figure 4 (b)
```{r colour selection}
library('RColorBrewer')
library('pheatmap')
my_color_annotation<-list(group = c(Control_D0 = "#FFFF08", 
                                    Infected_D2 = "#FFDA08",
                                    Infected_D4 ="#FFB508",
                                    Infected_D7 ="#FF8C08",
                                    Infected_D14 ="#FF5508"),
                          cluster=c('5'="#99E600",'1'="#68DC00",'2'="#00D43D",
                                    '7'="#00978C",'3'="#095E9F",'6'="#132AA9",'4'="#6D08A5"))
hm_color<- colorRampPalette(rev(brewer.pal(11, "RdBu")))(100)

HM_vsdmat<-pheatmap(vsdmat, scale = 'row', border_color = 'NA',
                    color = hm_color,show_rownames = FALSE,
                    cluster_rows = TRUE, cluster_cols = FALSE,
                    annotation_col=coldata_hm,
                    annotation_colors =my_color_annotation,
                    clustering_distance_rows = 'correlation',
                    angle_col =45)

clusterlist<-as.data.frame(vsdmat[HM_vsdmat$tree_row[['order']],]) #obtain gene clusters #each gene cluster is exported for functional enrichment analysis
clus<-as.data.frame(cutree(HM_vsdmat$tree_row, k=7))
colnames(clus)<-c('cluster')
clus<-clus[rownames(clusterlist),,drop=FALSE]
clusterlist$cluster<-clus$cluster
write.csv(clusterlist, "../Data/clusterlist.csv")
roldata_hm<-clusterlist[,c(25),drop=FALSE]
roldata_hm$cluster<-factor(roldata_hm$cluster,levels = c("5",'1','2','7','3','6','4'))

HM_vsdmat<-pheatmap(vsdmat, scale = 'row', border_color = 'NA',
                    color = hm_color,show_rownames = FALSE,
                    cluster_rows = TRUE, cluster_cols = FALSE,
                    annotation_col=coldata_hm,
                    annotation_row = roldata_hm,
                    annotation_colors =my_color_annotation,
                    clustering_distance_rows = 'correlation',
                    angle_col =45)
```

rmarkdown::render(rstudioapi::getSourceEditorContext()$path)

